\input{settings.tex}

\pgfplotsset{every tick label/.append style={font=\tiny}}

\begin{document}
  \title{Computer Haptics: Assignment 2}
  \maketitle

  \section*{Task 1 - Spring}

  For this experiment, the absolute position $x_{user}$ was multiplied with the stiffness factor $f$ to calculate the Force. The following graph shows the output when moving the handle in one direction. The displayed results were generated with $f = 0.002$.

  \begin{figure}[H]
    \centering
    \begin{tikzpicture}
      \begin{axis}[
        xlabel={$x_{user}$},
        ylabel={F},
        xmin=0, xmax=2300,
        ymin=0, ymax=4.50,
        xtick={500,1000,1500,2000},
        ytick={1,2,3,4},
        xtick pos=left,
        ytick pos=left,
      ]

        \addplot [ only marks, color=cyan ] table {res/measurements-2-1.txt};

      \end{axis}
    \end{tikzpicture}
  \end{figure}

  Increasing the stiffness resulted in the expected instability of the device, which started to show at $f = 0.008$. This can be noticed by the paddle bouncing from left to right without any interaction, due to imprecision of the \textit{Hapkit}. The high force moves the paddle beyond the zero position into the \textit{"opposite spring"}.

  \section*{Task 2 - Friction}

  Friction was rendered in two ways, coulomb and viscous. For \textit{Coulomb Friction}, force is applied against the movement direction of the handle in a constant way. Meaning no force is applied if velocity $v(t) = 0$. For a velocity $v(t) \neq 0$, the force is calculated by
  \begin{equation*}
    F_{coulomb} = -b * sgn(v(t))
  \end{equation*},
  where $b$ is the constant defining the friction. The following graph shows an experiment with coulomb friction, with $b = 1$:

  \begin{figure}[H]
    \centering
    \begin{tikzpicture}
      \begin{axis}[
        xlabel={$v(t)$},
        ylabel={F},
        xmin=-5, xmax=5,
        ymin=-2.5, ymax=2.5,
        xtick={-4,-2,0,2,4},
        ytick={-2,-1,0,1,2},
        xtick pos=left,
        ytick pos=left,
      ]

        \addplot [ only marks, color=cyan ] table {res/measurements-2-2-1.txt};

      \end{axis}
    \end{tikzpicture}
  \end{figure}

  \textit{Viscous Friction} on the other hand, acts similar if the velocity is zero, but renders and increasing friction for increasing velocity. This force is calculated as
  \begin{equation*}
    F_{viscous} = -b * v(t)
  \end{equation*}
  and has been used for the experiment visualized in the following graph, with $b = 0.75$:

  \begin{figure}[H]
    \centering
    \begin{tikzpicture}
      \begin{axis}[
        xlabel={$v(t)$},
        ylabel={F},
        xmin=-5, xmax=6,
        ymin=-5, ymax=5,
        xtick={-4,-2,0,2,4,5},
        ytick={-4,-2,0,2,4},
        xtick pos=left,
        ytick pos=left,
      ]

        \addplot [ very thick, color=cyan ] table {res/measurements-2-2-2.txt};

      \end{axis}
    \end{tikzpicture}
  \end{figure}

  Due to some imprecision, the velocity is only computed into force if $v(t) > 0.2$, as the device would become unstable otherwise. Therefore we get a bit more \textit{zero-values} for $F_{coulomb}$ and $F_{viscous}$ in the above graph.

  \section*{Task 3 - Virtual Wall}

  The virtual wall was rendered at position $pos_{wall} = 5mm$, which was calculated to an absolute position $x_{wall}$ at angle $\alpha \approx 4.76$. The force was calculated by multiplying the distance past the wall with the constant $k$:
  \begin{equation*}
    F = k * (x_{wall} - x_{user})
  \end{equation*}
  The following graph shows the applied force, when moving past the border of the wall, for $k = 0.75$.

  \begin{figure}[H]
    \centering
    \begin{tikzpicture}
      \begin{axis}[
        xlabel={$x_{user}$},
        ylabel={F},
        xmin=0, xmax=1000,
        ymin=-100, ymax=500,
        xtick={250,500,750},
        ytick={0,100,200,300,400},
        xtick pos=left,
        ytick pos=left,
      ]

        \addplot [ only marks, color=cyan ] table {res/measurements-2-3.txt};

        \addplot [ no marks, smooth, color=red ] coordinates { (340,-100) (340,500) } node[below] {$x_{wall}$};

      \end{axis}
    \end{tikzpicture}
  \end{figure}

  For $k > 1.25$, the device was found to be unstable as the resulting Force $F \approx 12.5$ created to much pressure on the rather fragile \textit{Hapkit}.

  \section*{Task 4 - Hard Surface}

  For this example, an additional vibration force, $F_{vibration}$, is rendered, when entering the wall. Time, $t$ in seconds, is measured upon entering the wall and used with the following damped sine wave equation to calculate the force to add to the constant wall force $F_{wall} = 100 N/m$:
  \begin{equation*}
    F_{vibration} = F_{wall} * e^{-t} * cos(2 * \pi * t)
  \end{equation*}
  The given time-steps, $t_{loop}$, in the graph below, represent the runs of the \texttt{loop()} function within the code.

  \begin{figure}[H]
    \centering
    \begin{tikzpicture}
      \pgfplotsset{set layers}

      \begin{axis} [
          scale only axis,
          xmin=0, xmax=1750,
          ymin=-250, ymax=250,
          xtick={0,500,1000,1500},
          ytick={-200,-150,-100,-50,0,50,100,150,200},
          axis y line*=left,
          xlabel={$t_{loop}$},
          ylabel style = {align=center},
          ylabel={$F$},
          legend style={at={(0.02,0.98)}, anchor=north west}
        ]

        \addplot [ no marks, color=cyan ] table {res/measurements-2-4-1.txt};
        \legend{$F$}
      \end{axis}

      \begin{axis} [
          scale only axis,
          xmin=0, xmax=1750,
          ymin=-2, ymax=2,
          ytick={-1,0,1},
          axis y line*=right,
          axis x line=none,
          ylabel style = {align=center},
          ylabel={$v$},
        ]

        \addplot [ no marks, color=orange ] table {res/measurements-2-4-2.txt};
        \legend{$v$}
      \end{axis}
      \end{tikzpicture}
  \end{figure}

  We can see the point at which the surface is hit by the paddle, as the motor applies force in a vibrating way. Those spikes in force occur, since we are directly at the border of the wall. During the time inside the wall, we can see the slight change in force until we leave the wall again at $t_{loop} \approx 1500$.

  \section*{Task 5 - Textures}

  To simulate textures, the force $F$ for velocity $v$ was calculated using a damper $b$ with the following formula, if the position is inside a damping area:
  \begin{equation*}
    F_{v} = -b_{v} * v_{t}
  \end{equation*}
  The damper itself was set to a value of 1 to minimize vibrations, that may occur due to the discontinuity in force. Width of the damping area $w = 0.01m$.

  \begin{figure}[H]
    \centering
    \begin{tikzpicture}
      \begin{axis}[
        xlabel={$x_{user}$},
        ylabel={F},
        xmin=0, xmax=3000,
        ymin=-2, ymax=2,
        xtick={0,500,1500,2500},
        ytick={-1,0,1},
        xtick pos=left,
        ytick pos=left,
      ]

        \addplot [ no marks, color=cyan ] table {res/measurements-2-5-1.txt};

        \addplot [ no marks, color=orange ] table {res/measurements-2-5-2.txt};

        \legend{$F_{v}$, $v$}

      \end{axis}
    \end{tikzpicture}
  \end{figure}

  The first texture part started at the absolute position $x_{user} = 0$. We can clearly see where the areas are located as the force there is zero. Since we are moving in a positive direction, we get a negative force, countering our velocity. The variations in force are to be expected, as we don't have a stable velocity $v$.

  \section*{Resources}

  Apart from the provided references on the assignment sheet, the example-code and instructions given at the website of the \textit{Hapkit} project,\\ \texttt{hapkit.stanford.edu}, were used to solve this assignment.

\end{document}