\input{settings.tex}

\pgfplotsset{every tick label/.append style={font=\tiny}}

\begin{document}
  \title{Computer Haptics: Assignment 1}
  \maketitle

  \section*{Task 1}

  While assembling the \textit{Hapkit}, the following important points were noticed:
  \begin{itemize}
    \item To get decent measurements of the magnetic sensor, the magnet should be as close as possible.
    \item It's best to wrap the string around the motor several times to create proper friction.
    \item There should be proper tension on the string, so the paddle does not move if the motor is not rotating.
  \end{itemize}

  \section*{Task 2}

  As an initial step, the values returned by the magnetic sensor of the \textit{Hapkit} were analysed to determine the highest one returned before flipping to zero. This value $m$ can be different for each device. In this case $m = 972$ was measured. Since the magnetic sensor records starting at zero after a full rotation, those flips, that occur after a full rotation, are represented by $f$. Using those variables, we can calculate the absolute position $p_{abs}$ with
  \begin{equation*}
    p_{abs} = \phi + (f * m)
  \end{equation*}
  where $\phi$ is the value returned by the magnetic sensor.
  To obtain the angle of the paddle $\alpha$, we need to find a constant to map from an absolute position. We can estimate that constant by calculating the average distance between the measurements from \textit{Task 3}. This leaves us with an average change $c = 77,13$ in absolute position every degree. Concluding from that, we get
  \begin{equation*}
    \alpha = \frac{\phi + (f * m)}{c} = \frac{p_{abs}}{c}
  \end{equation*}.
  This formula applies to a degree range starting at 0, as $p_{abs}$ is always a positive number, since we starting calibrating the paddle from the leftmost point. One would have to subtract 50 degrees to get a value between -50 and 50 for the \textit{Hapkit} device. Due to the imprecision of the device, the degree marks on the paddle, don't match the absolute position, as it moves from approximately -55 degrees to 55. To match $p_{abs} = 0$ to the leftmost mark on the paddle, for which $p_{abs} = 347$, we would have to subtract that from $p_{abs}$.

  \section*{Task 3}

  We can obtain the following graph by measuring the absolute positions at different degrees. The paddle itself is marked at every 10 degree step.

  \begin{figure}[H]
    \centering
    \begin{tikzpicture}
      \begin{axis}[
        xlabel={$p_{abs}$},
        ylabel={$\alpha$},
        xmin=0, xmax=8500,
        ymin=-55, ymax=55,
        xtick={1000,2000,3000,4000,5000,6000,7000,8000},
        ytick={-50,-25,0,25,50},
        xtick pos=left,
        ytick pos=left,
      ]

      \addplot[ no marks ] coordinates { (0,-55) (8500,55) };

      \addplot[ cyan, only marks ] coordinates {
        (347,-50) (1060,-40) (1980,-30) (2670,-20) (3400,-10) (4200,0) (4960,10) (5870,20) (6500,30) (7110,40) (8060,50)
      };

      \end{axis}
    \end{tikzpicture}
  \end{figure}
  
  The blue dots represent the measure values. We can clearly see some imprecision, which is due to the imperfection of the motor-mount and the string itself. Furthermore, the it's worth mentioning that the paddle moves from approximately -55 degrees to 55.

\end{document}